---
title: "Code for Sociodemographic Disparities in Exposures to Inorganic Contaminants in United States Public Water Systems"
authors: Mona Q. Dai, Xindi C. Hu, Brent A. Coull, Chris Campbell, David Q. Andrews, Olga V. Naidenko, Elsie M. Sunderland
output: html_document
last_updated: 9/6/24

Purpose: Run hurdle models and calculate odds ratios for effect sizes at PWS service area and county scales.
Note: Ensure working directory is set to where data is saved.
---

Clear environment.
```{r}
rm(list=ls())
gc()

```

Load libraries.
```{r}
library(tidyverse)
library(mltools)
library(tidymodels)
library(workflows)
library(caret)

library(mgcv)
library(lme4)
library(lmerTest)

```

Set working directory (path where code + data are saved).
```{r}
setwd('path')

```

Load functions.
```{r}
source('a_Functions.R')

```

Load data.
```{r}
source('b_Load_Data.R')

```

Run models.
```{r, include = FALSE}
source('c_Hurdle_PWS.R')
source('d_Hurdle_County.R')

```

======================================================================================
Create county OR table.
```{r}
# Skip urban_Mn_nonz_county (N/A)
county_OR <- CreateORtab(
  # Binomial
  As_binom_county, urban_As_binom_county, rural_As_binom_county, lg_As_binom_county, sm_As_binom_county,
  Cr_binom_county, urban_Cr_binom_county, rural_Cr_binom_county, lg_Cr_binom_county, sm_Cr_binom_county,
  Mn_binom_county, urban_Mn_binom_county, rural_Mn_binom_county, lg_Mn_binom_county, sm_Mn_binom_county,
  Se_binom_county, urban_Se_binom_county, rural_Se_binom_county, lg_Se_binom_county, sm_Se_binom_county,
  # Nonzero
  As_nonz_county, urban_As_nonz_county, rural_As_nonz_county, lg_As_nonz_county, sm_As_nonz_county,
  Cr_nonz_county, urban_Cr_nonz_county, rural_Cr_nonz_county, lg_Cr_nonz_county, sm_Cr_nonz_county,
  Mn_nonz_county, rural_Mn_nonz_county, rural_Mn_nonz_county, lg_Mn_nonz_county, sm_Mn_nonz_county,
  Se_nonz_county, urban_Se_nonz_county, rural_Se_nonz_county, lg_Se_nonz_county, sm_Se_nonz_county
)

```

View county OR table.
```{r}
# county_OR$binom_tab
# county_OR$nonz_tab
# 
# county_OR$As_binom_tab
# county_OR$As_nonz_tab
# county_OR$Cr_binom_tab
# county_OR$Cr_nonz_tab
# county_OR$Mn_binom_tab
# county_OR$Mn_nonz_tab
# county_OR$Se_binom_tab
# county_OR$Se_nonz_tab

```

======================================================================================
Create PWS OR table.
```{r}
pws_OR <- CreateORtab(
  # Binomial
  As_binom_pws, urban_As_binom_pws, rural_As_binom_pws, lg_As_binom_pws, sm_As_binom_pws,
  Cr_binom_pws, urban_Cr_binom_pws, rural_Cr_binom_pws, lg_Cr_binom_pws, sm_Cr_binom_pws,
  Mn_binom_pws, urban_Mn_binom_pws, rural_Mn_binom_pws, lg_Mn_binom_pws, sm_Mn_binom_pws,
  Se_binom_pws, urban_Se_binom_pws, rural_Se_binom_pws, lg_Se_binom_pws, sm_Se_binom_pws,
  # Nonzero
  As_nonz_pws, urban_As_nonz_pws, rural_As_nonz_pws, lg_As_nonz_pws, sm_As_nonz_pws,
  Cr_nonz_pws, urban_Cr_nonz_pws, rural_Cr_nonz_pws, lg_Cr_nonz_pws, sm_Cr_nonz_pws,
  Mn_nonz_pws, urban_Mn_nonz_pws, rural_Mn_nonz_pws, lg_Mn_nonz_pws, sm_Mn_nonz_pws,
  Se_nonz_pws, urban_Se_nonz_pws, rural_Se_nonz_pws, lg_Se_nonz_pws, sm_Se_nonz_pws
)

```

View PWS OR table.
```{r}
# pws_OR$binom_tab
# pws_OR$nonz_tab
# 
# pws_OR$As_binom_tab
# pws_OR$As_nonz_tab
# pws_OR$Cr_binom_tab
# pws_OR$Cr_nonz_tab
# pws_OR$Mn_binom_tab
# pws_OR$Mn_nonz_tab
# pws_OR$Se_binom_tab
# pws_OR$Se_nonz_tab

```


========================================================================================================
Calculate confidence intervals by urban/rural & large/small PWS.
```{r, warning=F, include=F}
source('e_Confidence_Intervals.R')

```

Adjust PWS & county df for 1 standard deviation increase in PWS sociodemographic factors.
```{r}
# PWS
As_PWS_binom_df1 <- AdjCounty(As_PWS_binom_df0, uspwsAs0, uspwsAs0)
Cr_PWS_binom_df1 <- AdjCounty(Cr_PWS_binom_df0, uspwsCr0, uspwsCr0)
Mn_PWS_binom_df1 <- AdjCounty(Mn_PWS_binom_df0, uspwsMn0, uspwsMn0)
Se_PWS_binom_df1 <- AdjCounty(Se_PWS_binom_df0, uspwsSe0, uspwsSe0)

As_PWS_nonz_df1 <- AdjCounty(As_PWS_nonz_df0, filter(uspwsAs0, NONZERO == 1), uspwsAs0)
Cr_PWS_nonz_df1 <- AdjCounty(Cr_PWS_nonz_df0, filter(uspwsCr0, NONZERO == 1), uspwsCr0)
Mn_PWS_nonz_df1 <- AdjCounty(Mn_PWS_nonz_df0, filter(uspwsMn0, NONZERO == 1), uspwsMn0)
Se_PWS_nonz_df1 <- AdjCounty(Se_PWS_nonz_df0, filter(uspwsSe0, NONZERO == 1), uspwsSe0)

# County
As_County_binom_df1 <- AdjCounty(As_County_binom_df0, uscountyAs0, uspwsAs0)
Cr_County_binom_df1 <- AdjCounty(Cr_County_binom_df0, uscountyCr0, uspwsCr0)
Mn_County_binom_df1 <- AdjCounty(Mn_County_binom_df0, uscountyMn0, uspwsMn0)
Se_County_binom_df1 <- AdjCounty(Se_County_binom_df0, uscountySe0, uspwsSe0)

As_County_nonz_df1 <- AdjCounty(As_County_nonz_df0, filter(uscountyAs0, NONZERO == 1), uspwsAs0)
Cr_County_nonz_df1 <- AdjCounty(Cr_County_nonz_df0, filter(uscountyCr0, NONZERO == 1), uspwsCr0)
Mn_County_nonz_df1 <- AdjCounty(Mn_County_nonz_df0, filter(uscountyMn0, NONZERO == 1), uspwsMn0)
Se_County_nonz_df1 <- AdjCounty(Se_County_nonz_df0, filter(uscountySe0, NONZERO == 1), uspwsSe0)

```

PWS.
```{r}
## Filter for predictors of primary interest: limited English-speaking house holds (NOENG), Black residents (BLACK), & PIR<2 (POVERTY200)
# Binomial
as_burls_pws          <- filter(As_PWS_binom_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
as_burls_pws$Chemical <- 'Arsenic'
as_burls_pws$Hurdle   <- 'Binomial'
cr_burls_pws          <- filter(Cr_PWS_binom_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
cr_burls_pws$Chemical <- 'Chromium'
cr_burls_pws$Hurdle   <- 'Binomial'
mn_burls_pws          <- filter(Mn_PWS_binom_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
mn_burls_pws$Chemical <- 'Manganese'
mn_burls_pws$Hurdle   <- 'Binomial'
se_burls_pws          <- filter(Se_PWS_binom_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
se_burls_pws$Chemical <- 'Selenium'
se_burls_pws$Hurdle   <- 'Binomial'
# Nonzero regression
as_nurls_pws          <- filter(As_PWS_nonz_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
as_nurls_pws$Chemical <- 'Arsenic'
as_nurls_pws$Hurdle   <- 'Nonzero'
cr_nurls_pws          <- filter(Cr_PWS_nonz_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
cr_nurls_pws$Chemical <- 'Chromium'
cr_nurls_pws$Hurdle   <- 'Nonzero'
mn_nurls_pws          <- filter(Mn_PWS_nonz_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
mn_nurls_pws$Chemical <- 'Manganese'
mn_nurls_pws$Hurdle   <- 'Nonzero'
se_nurls_pws          <- filter(Se_PWS_nonz_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
se_nurls_pws$Chemical <- 'Selenium'
se_nurls_pws$Hurdle   <- 'Nonzero'

## Combine all
all_bnurls_pws <- rbind(as_burls_pws, cr_burls_pws, mn_burls_pws, se_burls_pws,
                        as_nurls_pws, cr_nurls_pws, mn_nurls_pws, se_nurls_pws)

## Resort by predictor
all_bnurls_pws <- Orderby(all_bnurls_pws, 'Predictor', de.creasing = F)

## Break back down into individual df
urban_burls_pws <- filter(all_bnurls_pws, ID == "Urban" & Hurdle == "Binomial")
rural_burls_pws <- filter(all_bnurls_pws, ID == "Rural" & Hurdle == "Binomial")
lg_burls_pws    <- filter(all_bnurls_pws, ID == "Large PWS" & Hurdle == "Binomial")
sm_burls_pws    <- filter(all_bnurls_pws, ID == "Small PWS" & Hurdle == "Binomial")

urban_nurls_pws <- filter(all_bnurls_pws, ID == "Urban" & Hurdle == "Nonzero")
rural_nurls_pws <- filter(all_bnurls_pws, ID == "Rural" & Hurdle == "Nonzero")
lg_nurls_pws    <- filter(all_bnurls_pws, ID == "Large PWS" & Hurdle == "Nonzero")
sm_nurls_pws    <- filter(all_bnurls_pws, ID == "Small PWS" & Hurdle == "Nonzero")

```

```{r}
## Filter for predictors of primary interest
# Binomial
as_burls_county          <- filter(As_County_binom_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
as_burls_county$Chemical <- 'Arsenic'
as_burls_county$Hurdle   <- 'Binomial'
cr_burls_county          <- filter(Cr_County_binom_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
cr_burls_county$Chemical <- 'Chromium'
cr_burls_county$Hurdle   <- 'Binomial'
mn_burls_county          <- filter(Mn_County_binom_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
mn_burls_county$Chemical <- 'Manganese'
mn_burls_county$Hurdle   <- 'Binomial'
se_burls_county          <- filter(Se_County_binom_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
se_burls_county$Chemical <- 'Selenium'
se_burls_county$Hurdle   <- 'Binomial'
# Nonzero regression
as_nurls_county          <- filter(As_County_nonz_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
as_nurls_county$Chemical <- 'Arsenic'
as_nurls_county$Hurdle   <- 'Nonzero'
cr_nurls_county          <- filter(Cr_County_nonz_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
cr_nurls_county$Chemical <- 'Chromium'
cr_nurls_county$Hurdle   <- 'Nonzero'
mn_nurls_county          <- filter(Mn_County_nonz_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
mn_nurls_county$Chemical <- 'Manganese'
mn_nurls_county$Hurdle   <- 'Nonzero'
se_nurls_county          <- filter(Se_County_nonz_df1, Predictor %in% c('NOENG', 'BLACK', 'POVERTY200'))
se_nurls_county$Chemical <- 'Selenium'
se_nurls_county$Hurdle   <- 'Nonzero'

## Combine all
all_bnurls_county <- rbind(as_burls_county, cr_burls_county, mn_burls_county, se_burls_county,
                           as_nurls_county, cr_nurls_county, mn_nurls_county, se_nurls_county)

## Resort by predictor
all_bnurls_county <- Orderby(all_bnurls_county, 'Predictor', de.creasing = F)

## Break back down into individual df
urban_burls_county <- filter(all_bnurls_county, ID == "Urban" & Hurdle == "Binomial")
rural_burls_county <- filter(all_bnurls_county, ID == "Rural" & Hurdle == "Binomial")
lg_burls_county    <- filter(all_bnurls_county, ID == "Large PWS" & Hurdle == "Binomial")
sm_burls_county    <- filter(all_bnurls_county, ID == "Small PWS" & Hurdle == "Binomial")

urban_nurls_county <- filter(all_bnurls_county, ID == "Urban" & Hurdle == "Nonzero")
rural_nurls_county <- filter(all_bnurls_county, ID == "Rural" & Hurdle == "Nonzero")
lg_nurls_county    <- filter(all_bnurls_county, ID == "Large PWS" & Hurdle == "Nonzero")
sm_nurls_county    <- filter(all_bnurls_county, ID == "Small PWS" & Hurdle == "Nonzero")

```

-------------------------
Join adjusted PWS & County dataframes into 1 dataframe.
```{r}
As_PWScounty_binom_df1 <- JoinAdjusted(As_PWS_binom_df1, As_County_binom_df1)
Cr_PWScounty_binom_df1 <- JoinAdjusted(Cr_PWS_binom_df1, Cr_County_binom_df1)
Mn_PWScounty_binom_df1 <- JoinAdjusted(Mn_PWS_binom_df1, Mn_County_binom_df1)
Se_PWScounty_binom_df1 <- JoinAdjusted(Se_PWS_binom_df1, Se_County_binom_df1)

As_PWScounty_nonz_df1 <- JoinAdjusted(As_PWS_nonz_df1, As_County_nonz_df1)
Cr_PWScounty_nonz_df1 <- JoinAdjusted(Cr_PWS_nonz_df1, Cr_County_nonz_df1)
Mn_PWScounty_nonz_df1 <- JoinAdjusted(Mn_PWS_nonz_df1, Mn_County_nonz_df1)
Se_PWScounty_nonz_df1 <- JoinAdjusted(Se_PWS_nonz_df1, Se_County_nonz_df1)

```









